---
title: "Project_430"
author: "Kashyap Ava"
date: "2025-02-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading Data

```{r}
data_biomet <- read.csv("/Users/kashyapava/Desktop/Crop Sciences/New Data/eddypro_MaizeControl_2023_01-08_biomet_2023-09-18T163846_adv.csv", header = TRUE)

# Remove the rows with units in it
units_biomet <- data_biomet[1,]
data_biomet <- data_biomet[-1,]

head(data_biomet)
```

```{r}
units_biomet
```



```{r}
data_full_output <- read.csv("/Users/kashyapava/Desktop/Crop Sciences/New Data/eddypro_MaizeControl_2023_01-08_full_output_2023-09-18T163846_adv.csv", header = TRUE)

# Select the row to be used as the header (e.g., the first row)
new_header <- as.character(unlist(data_full_output[1, ]))

# Remove the selected row from the data frame
data_full_output <- data_full_output[-1, ]

# Assign the new header to the data frame
colnames(data_full_output) <- new_header

# Remove the row with units
units_full_output <- data_full_output[1,]
data_full_output <- data_full_output[-1,]

# View the modified data frame
head(data_full_output)
```

```{r}
units_full_output
```


## Converting the string type numerical values to float

```{r}
# Exclude the 'date' and 'time' columns
data_numeric <- data_biomet[, !(names(data_biomet) %in% c("date", "time"))]

# Convert the remaining columns to numeric
data_numeric[] <- lapply(data_numeric, function(x) as.numeric(as.character(x)))

# Combine the 'date' and 'time' columns back if needed
data_biomet_clean <- cbind(data_biomet[, c("date", "time")], data_numeric)

# View the modified data frame
head(data_biomet_clean)
```

## Data Cleaning

```{r}
data_biomet_clean$SHF <- apply(data_biomet_clean[, c("SHF_1_1_1", "SHF_1_1_2")], 1, function(row) {
  valid_values <- row[row != -9999]  # Filter out the -9999 values
  if (length(valid_values) == 0) {
    return(-9999)  # If all values are -9999, return -9999
  } else {
    return(mean(valid_values))  # Otherwise, return the mean of the valid values
  }
})

data_biomet_clean$Ts <- apply(data_biomet_clean[, c("Ts_1_1_1", "Ts_1_1_1.1", "Ts_1_2_1", "Ts_1_3_1", "Ts_1_4_1", "Ts_2_1_1", "Ts_2_2_1", "Ts_2_3_1", "Ts_2_4_1", "Ts_3_1_1", "Ts_3_2_1", "Ts_3_3_1", "Ts_3_4_1", "Ts_3_5_1", "Ts_3_6_1", "Ts_3_7_1", "Ts_3_8_1", "Ts_3_9_1", "Ts_4_1_1", "Ts_4_2_1", "Ts_4_3_1", "Ts_4_4_1", "Ts_4_5_1", "Ts_4_6_1", "Ts_4_7_1", "Ts_4_8_1", "Ts_4_9_1")], 1, function(row) {
  valid_values <- row[row != -9999]  # Filter out the -9999 values
  if (length(valid_values) == 0) {
    return(-9999)  # If all values are -9999, return -9999
  } else {
    return(mean(valid_values))  # Otherwise, return the mean of the valid values
  }
})

data_biomet_clean$SWC <- apply(data_biomet_clean[, c("SWC_1_1_1", "SWC_1_2_1", "SWC_1_3_1", "SWC_1_4_1", "SWC_2_1_1", "SWC_2_2_1", "SWC_2_3_1", "SWC_2_4_1", "SWC_3_1_1", "SWC_3_2_1", "SWC_3_3_1", "SWC_3_4_1", "SWC_3_5_1", "SWC_3_6_1", "SWC_3_7_1", "SWC_3_8_1", "SWC_3_9_1", "SWC_4_1_1", "SWC_4_2_1", "SWC_4_3_1", "SWC_4_4_1", "SWC_4_5_1", "SWC_4_6_1", "SWC_4_7_1", "SWC_4_8_1", "SWC_4_9_1")], 1, function(row) {
  valid_values <- row[row != -9999]  # Filter out the -9999 values
  if (length(valid_values) == 0) {
    return(-9999)  # If all values are -9999, return -9999
  } else {
    return(mean(valid_values))  # Otherwise, return the mean of the valid values
  }
})

head(data_biomet_clean)

```


## Forming a regression dataset for the time series analysis

```{r}
regression_data <- data_biomet_clean[, c("date","time", "DOY", "SWin", "SWout", "LWin", "LWout", "Rr", "Rn", "Ta_1_1_1", "RH", "Tc_1_1_1", "PPFD", "PPFDr", "P_rain", "SHF", "Ts", "SWC")]

head(regression_data)
```


# Extracting the target variable - CH4 flux

```{r}
data_full_output <- data_full_output[, 1:159]
data_full_output_1 <- data_full_output[, !(names(data_full_output) %in% c("filename", "date", "time" ))]

# Convert the remaining columns to numeric
data_full_output_1[] <- lapply(data_full_output_1, function(x) as.numeric(as.character(x)))

ch4_flux <- data_full_output_1[,"ch4_flux"]
ch4_fname <- data_full_output[, "filename"]

new_predictors <- data_full_output[, (names(data_full_output) %in% c("w/ch4_cov", "T*", "LE", "h2o_flux", "ET", "w/h2o_cov",  "w/co2_cov", "co2_flux", "(z-d)/L", "v_var"))]

data_reg <- cbind(regression_data,ch4_flux)
data_reg_1 <- cbind(data_reg, ch4_fname)
data_reg_1 <- cbind(data_reg_1, new_predictors)
head(data_reg_1)
```

```{r}
# Load necessary library
library(dplyr)

# Replace -9999 with NA
data_full_output_1[data_full_output_1 == -9999] <- NA

# Remove columns with more than 50% NA values
threshold <- 0.5 * nrow(data_full_output_1)
data_clean <- data_full_output_1[, colSums(is.na(data_full_output_1)) <= threshold]

# Remove rows with any NA values
data_clean <- na.omit(data_clean)

dim(data_clean)
```

```{r}
head(data_clean)
```


```{r}
colnames(data_clean)
```


```{r}
# Remove unwanted columns
excluded_columns <- c("DOY", "daytime", "file_records", "used_records")  
numeric_data <- data_clean %>% select(-all_of(excluded_columns)) %>% select(where(is.numeric))

# Compute Spearman correlation with ch4_flux
correlations <- suppressWarnings(sapply(numeric_data, function(col) cor(col, data_clean$ch4_flux, method = "spearman")))

# Sort by absolute correlation value in descending order
sorted_correlations <- sort(abs(correlations), decreasing = TRUE)

filtered_correlations <- sorted_correlations[abs(sorted_correlations) > 0.3]

# Print results
print(filtered_correlations)
```

```{r}
as.vector.data.frame(filtered_correlations)
```



Explore these predictors to include them in the model:

un_ch4_flux:  Uncorrected Methane flux      

w/ch4_cov : Covariance between w and variable ch4               

T* : Scaling temperature (K)               

LE : Corrected latent heat flux (W m-2)            

un_LE : Uncorrected latent heat flux (W m-2)

h2o_flux : Water flux              

ET : Evapotranspiration flux (mm hour-1)      

un_h2o_flux : Uncorrected Water flux    

w/h2o_cov : Covariance between w and variable h2o      

un_co2_flux : Uncorrected Carbon dioxide flux        

w/co2_cov : Covariance between w and variable co2
        
co2_flux : Carbon dioxide flux         

(z-d)/L : Monin-Obukhov stability parameter            

v_var : variance of the v wind component


Out of intuition and with caution these were selected to be included:

w/ch4_cov : Covariance between w and variable ch4               

T* : Scaling temperature (K)               

LE : Corrected latent heat flux (W m-2)

h2o_flux : Water flux              

ET : Evapotranspiration flux (mm hour-1)   

w/h2o_cov : Covariance between w and variable h2o      

w/co2_cov : Covariance between w and variable co2
        
co2_flux : Carbon dioxide flux         

(z-d)/L : Monin-Obukhov stability parameter            

v_var : variance of the v wind component




```{r}
# Utilizing the filename column to remove rows with "not enough data"
# Remove rows where filename is "not enough data"

data_reg_clean <- data_reg_1[data_reg_1$ch4_fname != "not_enough_data", ]
head(data_reg_clean)
```


```{r}
# Renaming them as the predictor names
data_reg_clean$FCH4 <- data_reg_clean[, c("ch4_flux")]

data_reg_clean$Ta <- data_reg_clean[, c("Ta_1_1_1")]

data_reg_clean$Tc <- data_reg_clean[, c("Tc_1_1_1")]

data_final_clean <- data_reg_clean[, !(names(data_reg_clean) %in% c("Ta_1_1_1", "Tc_1_1_1", "ch4_flux", "ch4_fname" ))]

head(data_final_clean)
```

```{r}
# Count the number of rows with values equal to -9999 in each column
counts <- sapply(data_final_clean, function(x) sum(x == -9999))

# Display the counts
print(counts)
```

# Checking if the null rows for predictors are common

```{r}
# Remove rows with NA in 'SWin'
df_cleaned <- data_final_clean[data_final_clean$RH != -9999, ]

# Count the number of rows with values equal to -9999 in each column
counts <- sapply(df_cleaned, function(x) sum(x == -9999))

# Display the counts
print(counts)
```

```{r}
# Remove rows with NA in 'SWin'
df_cleaned_x <- df_cleaned[df_cleaned$FCH4 != -9999, ]

# Count the number of rows with values equal to -9999 in each column
counts <- sapply(df_cleaned_x, function(x) sum(x == -9999))

# Display the counts
print(counts)
```

```{r}
# Remove rows with NA in 'SWin'
df_cleaned_y <- df_cleaned_x[df_cleaned_x$LE != -9999, ]

# Count the number of rows with values equal to -9999 in each column
counts <- sapply(df_cleaned_y, function(x) sum(x == -9999))

# Display the counts
print(counts)
```

```{r}
# Remove rows with NA in 'SWin'
df_cleaned_z <- df_cleaned_y[df_cleaned_y$co2_flux != -9999, ]

# Count the number of rows with values equal to -9999 in each column
counts <- sapply(df_cleaned_z, function(x) sum(x == -9999))

# Display the counts
print(counts)
```



```{r}
dim(df_cleaned_z)
```

```{r}
# Count the number of rows with values equal to -9999 in each column
counts <- sapply(df_cleaned_z, function(x) sum(x == -9999))

# Display the counts
print(counts)
```

No null values in the df_cleaned_z data set.




## Feature Engineering



## Cleaned data

```{r}
tsa_data <- df_cleaned_z

head(tsa_data)
```

Checking for null values:

```{r}
# Count the number of rows with values equal to -9999 in each column
counts <- sapply(tsa_data, function(x) sum(is.na(x)))

# Display the counts
print(counts)
```


```{r}
# Count the number of rows with values equal to -9999 in each column
counts <- sapply(tsa_data, function(x) sum(x == -9999))

# Display the counts
print(counts)
```

Dimension of the final data set for time series analysis:

```{r}
dim(tsa_data)
```


## Exploratory Data Analysis

```{r}
tsa_data_stored <- tsa_data
```

```{r}
# Replace -9999 values in FCH4 column with NA
tsa_data_stored$FCH4[tsa_data_stored$FCH4 == -9999] <- NA

# Combine date and time columns into a single POSIXct datetime column
tsa_data_stored$datetime <- as.POSIXct(paste(tsa_data_stored$date, tsa_data_stored$time), format="%Y-%m-%d %H:%M")

# Plot the data
plot(tsa_data_stored$datetime, tsa_data_stored$FCH4, type = "l",
     main = "Time Series Plot of FCH4",
     xlab = "Datetime",
     ylab = "FCH4",
     col = "blue",
     na.rm = TRUE)
```


From: 2023-05-11 12:00:00

To: 2023-07-10 05:00

For the STAT 430 project, has some null values in between but can be taken care of.



## Data Exploration for the project

From: 2023-05-11 15:00:00

To: 2023-07-10 05:00

Total rows - 2741 (after cleaning)

```{r}
proj_data <- tsa_data_stored[tsa_data_stored$datetime >= "2023-05-11 15:00:00" & tsa_data_stored$datetime <= "2023-07-10 05:00:00", ]

head(proj_data)
```

```{r}
dim(proj_data)
```

From: 2023-05-11 15:00:00

To: 2023-07-10 01:30

Total rows - 2726 (after cleaning with added predictors)

```{r}
sum(is.na(proj_data))
```


```{r}
# Plot the data
plot(proj_data$datetime, proj_data$FCH4, type = "l",
     main = "Time Series Plot of FCH4",
     xlab = "Datetime",
     ylab = "FCH4 (µmol+1s-1m-2)",
     col = "blue",
     na.rm = TRUE)
```

Changed start date to 2023-05-11 15:00:00.

Total rows - 2726.


```{r}
summary(proj_data[, "P_rain"])
```


```{r}
# change the precipitation units from m to mm
proj_data[, "P_rain"] <- 1000 * proj_data[, "P_rain"]
summary(proj_data[, "P_rain"])
```

```{r}
# Define your data as named vectors
variable_names <- c("ch4_flux", "un_ch4_flux", "w/ch4_cov", "T*", "LE", 
                    "un_LE", "h2o_flux", "ET", "un_h2o_flux", "w/h2o_cov",
                    "un_co2_flux", "w/co2_cov", "co2_flux", "(z-d)/L", "v_var")

values <- c(1.0000000, 0.6289260, 0.6289260, 0.3280700, 0.3272453, 
            0.3270026, 0.3262913, 0.3262911, 0.3259612, 0.3259612, 
            0.3156560, 0.3156560, 0.3054115, 0.3033668, 0.3012782)

# Create a DataFrame
output_table <- data.frame(Variable = variable_names, Value = values)

# Print as a table
print(output_table, row.names = FALSE)
```

```{r}
# Save as CSV
write.csv(proj_data, "proj2_data_clean.csv", row.names = FALSE)
```








